---
title: 'Accessible metrics of access: Novel tools to measure immunization coverage, Phase 3 Initial Data Summary'
author: "Paul L. Delamater"
date: "1/27/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(kableExtra)
library(tmap)
library(RColorBrewer)
tmap_mode("view")
```

```{r read_clean_data, echo=FALSE, warning=FALSE, message=FALSE}

## CODE FROM  read_clean_phase3_data.R
## Minus plots and writing data out

## Read data
dat <- read_csv("../../data/BoyceGatesVaccineAcc_DATA_2021-04-20_0744.csv")

## Get initial number of observations
n_obs_original <- nrow(dat)

## Initial scan of position data shows a lot of issues...
##   Read in as character data
##   Let's use hh_lat and hh_long if available
##   Fill in missing from lat_tab and long_tab
##     But, note that lat_tab and long_tab have some reversed

## One entry has double ".."?
dat$lat_tab <- dat$lat_tab %>% str_replace("\\.\\.", "\\.") %>% as.numeric()

##
## Create single lat/long ----
##

## Make holder column for corrected *_tab versions
dat$lat_tab_upd <- dat$lat_tab
dat$long_tab_upd <- dat$long_tab

## Identify obvious errors
lat_long_flip <- which(dat$lat_tab > 1)
n_fix_flip <- length(lat_long_flip)

## Replace
dat$lat_tab[lat_long_flip] <- dat$long_tab_upd[lat_long_flip]
dat$long_tab[lat_long_flip] <- dat$lat_tab_upd[lat_long_flip]

## Create new field
dat$LAT <- dat$hh_lat
dat$LONG <- dat$hh_long

## Replace missing with "tablet" verions
dat$LAT[which(is.na(dat$LAT))] <- dat$lat_tab[which(is.na(dat$LAT))]
dat$LONG[which(is.na(dat$LONG))] <- dat$long_tab[which(is.na(dat$LONG))]

## Looks like one is a transcription error 0.13xxxx instead of 0.31xxxx
##    but, we cannot be sure, unfortunately
## Looks like one is simply an error (missing LONG, both entries are LAT)
## Looks like one is transciption error begins LONG, ends with LAT
## So, we will remove them for now
dat.na <- dat %>% filter(is.na(LAT) | is.na(LONG)) %>% nrow()
dat.errors <- dat %>% filter(LAT < 0.2 | LONG < 25 | LONG > 30.2)

## Remove temp columns
dat.errors <- nrow(dat.errors)

## Read cleaned version of data
dat <- read_csv("../../output/BoyceGatesVaccineAcc_DATA_2021-04-20_0744_clean_geo.csv")

## Get initial number of observations
n_obs_final <- nrow(dat)
```

## Observations

There are `r format(n_obs_original, big.mark = ",")` observations in the data. Of these, `r n_fix_flip` appeared to have reversed latitude and longitude coordinate values (which were corrected). `r dat.errors` observations appeared to have clearly erroneous coordinate values, and `r dat.na` observation was missing one or both coordinate values. These observations were removed. After removal of the observations with missing or erroneous coordinates, there are **`r format(n_obs_final, big.mark = ",")`** remaining (mapped below). 

```{r point_map, echo=FALSE, warning=FALSE, message=FALSE}
##
## Create spatial object
##
dat <- st_as_sf(dat %>% as.data.frame(),
                coords = c("LONG", "LAT"),
                crs = 4326)

basemap <- "CartoDB.Voyager"

## Map
tm_shape(dat) +
  tm_basemap(basemap) +
  tm_dots(size = 0.01) 

```

```{r parish_summary, echo=FALSE, warning=FALSE, message=FALSE}

##
## Evaluate the number of points with correct parish
##

## Read in parish boundaries
ug.par.sf <- read_sf("../../data/Uganda_parish_bounds_2010/Uganda_parish_bounds_2010.shp")

## Subset to parishes in the Kasese district
# kasese.sf <- ug.par.sf %>% filter(DNAME_2010 == "KASESE")

## Reproject parish bounds to WGS
kasese.sf <- st_transform(ug.par.sf,
                          crs = 4326)

## Spatial join points to polygons
## Note: only the name column of parish
dat <- st_join(dat, kasese.sf[,2])

## Rename column
names(dat)[ncol(dat)-1] <- "PARISH_GIS"

## Get Parish Codes/Names file
## Copy/pasted to create manually
par.codes <- read_csv("../../data/field_codes/ph1_parish_codes.csv")

## Join to data
dat <- merge(dat,
             par.codes,
             by.x = "parish",
             by.y = "ph1_parish",
             all.x = TRUE)

## How many Parishes don't match and do match?
n.parish.match <- sum(dat$PARISH_GIS == dat$`Parish of Residence`)
n.parish.mismatch <- sum(dat$PARISH_GIS != dat$`Parish of Residence`)

## Make variable
dat$PARISH_M <- rep(0, nrow(dat))
dat$PARISH_M[which(dat$PARISH_GIS == dat$`Parish of Residence`)] <- 1

## Summarize
par.match.tab <- st_drop_geometry(dat)[,c("Parish of Residence", "PARISH_M")] %>% 
  group_by(`Parish of Residence`) %>% 
  summarize_all(sum)
par.all.tab <- table(dat$`Parish of Residence`, useNA = "ifany")
par.match.tab$TOTAL <- par.all.tab

## Percent agreement
par.match.tab$PCT <- round(100 * par.match.tab$PARISH_M / par.match.tab$TOTAL, 1)

## Make pretty
pmt.d <- par.match.tab
pmt.d <- pmt.d[,c(1,3,2,4)]
names(pmt.d)[1:4] <- c("Reported Parish",
                       "Observations",
                       "Matches",
                       "Match Percent")
pmt.d$`Reported Parish`[6] <- "Missing"
```



## Observations by Parish

The observations were converted to a spatial data format using the GPS coordinates of the residence. This conversion allowed the data to be overlain with a spatial data layer containing the boundaries of the Uganda's parishes. The parish recorded at the visit was compared to the GIS-derived parish, and the agreement between the two are presented below.

```{r parish_table, echo=FALSE, warning=FALSE, message=FALSE}
kable(pmt.d,
      align = 'lccc') %>% 
  kable_styling(bootstrap_options = c("striped", 
                                      "hover", 
                                      "condensed", 
                                      "responsive"), 
                full_width = F)
```


## Observations by Village

SPACE HOLDER FOR OBSERVATIONS BY VILLAGE

```{r village_summary, echo=FALSE, warning=FALSE, message=FALSE}

```

```{r village_table, echo=FALSE, warning=FALSE, message=FALSE}

```


## Immunization Summary

```{r imm_card_summary, echo=FALSE, warning=FALSE, message=FALSE}
## Get number of observations with card
card_summary <- table(dat$card, useNA = "ifany")

## Make subset table of obs with card
dat_card <- dat %>% filter(card == 1)

## Get summary of each vaccine
vacc_summary <- lapply(st_drop_geometry(dat_card)[,62:69], table, useNA = "ifany") %>%
  bind_rows() %>% t() %>% as.data.frame()

## Make pretty
vacc_summary$Vaccine <- c("Bacillus Calmette–Guérin (BCG)",
                          "Polio (OPV)",
                          "Pentavalent (DPT-HepB-Hib), Dose #1",
                          "Pentavalent (DPT-HepB-Hib), Dose #2",
                          "Pentavalent (DPT-HepB-Hib), Dose #3",
                          "Measles",
                          "Rotavirus, Dose #1",
                          "Rotavirus, Dose #2")
names(vacc_summary)[1:3] <- c("Not Administered", "Administered", "Unknown")
vacc_summary$`Covered \\% (known)` <- round(100 * vacc_summary$Administered / nrow(dat_card), 2)
vacc_summary <- vacc_summary[,c(4,2,1,3,5)]
rownames(vacc_summary) <- NULL
```

```{r imm_no_card_summary, echo=FALSE, warning=FALSE, message=FALSE}
## Make subset table of obs with card
dat_no_card <- dat %>% filter(card == 0)

## Get summary of each vaccine
vacc_summary_nc <- table(dat_no_card$vacc_recall, useNA = "ifany")

```

Of the `r format(n_obs_final, big.mark = ",")` observations, **`r format(card_summary[2], big.mark = ",")`** had their vaccination card and **`r format(card_summary[1], big.mark = ",")`** did not have their vaccination card. `r format(card_summary[3], big.mark = ",")` of the observations were missing this information. This information is mapped below.

```{r imm_card_map, echo=FALSE, warning=FALSE, message=FALSE}
tm_shape(dat) +
  tm_basemap(basemap) +
  tm_dots(col = "card",
          style = "cat",
          palette = "Pastel1",
          colorNA = "black",
          legend.show = FALSE) +
  tm_add_legend(type = "fill",
                title = "Immunization Card",
                labels = c("Not Available", 
                           "Available",
                           "No Data"),
                col = c(brewer.pal(3, "Pastel1")[1:2], "black"))
```

### Specific Vaccines Summary

For each specific vaccine (and dose, when appropriate), we calculated the percent coverage (only for children with a card). Some observations were missing data, which we treated as "not" covered in the estimates below. Overall, the coverage percent (in this group) was quite high.

```{r utd_table, echo=FALSE, warning=FALSE, message=FALSE}
kable(vacc_summary,
      align = 'lcccc') %>% 
  kable_styling(bootstrap_options = c("striped", 
                                      "hover", 
                                      "condensed", 
                                      "responsive"), 
                full_width = F)
```

### Immunization Status Recall

For children without an immunization card, the respondent was asked to recall their vaccination status, with three potential responses: no vaccinations, most vaccinations, all vaccinations. Out of the `r format(card_summary[1], big.mark = ",")` observations with no card, `r vacc_summary_nc[1]` (`r round(100 * vacc_summary_nc[1] / card_summary[1], 2)`%) reported the child had received no vaccinations, `r vacc_summary_nc[2]` (`r round(100 * vacc_summary_nc[2] / card_summary[1], 2)`%) reported most vaccinations, and `r vacc_summary_nc[3]` (`r round(100 * vacc_summary_nc[3] / card_summary[1], 2)`%) reported all vaccinations (`r vacc_summary_nc[4]` had a missing entry). 


```{r imm_no_card_map, echo=FALSE, warning=FALSE, message=FALSE}
tm_shape(dat_no_card) +
  tm_basemap(basemap) +
  tm_dots(col = "vacc_recall",
          style = "cat",
          palette = "RdYlBu",
          colorNA = "black",
          legend.show = FALSE) +
  tm_add_legend(type = "fill",
                title = "Recall Status",
                labels = c("No Vaccinations", 
                           "Most Vaccinations",
                           "All Vaccinations",
                           "No Data"),
                col = c(brewer.pal(3, "RdYlBu"), "black"))
```